---
output:
  html_document:
    variant: markdown_github
    keep_md: yes
  md_document:
    variant: markdown_github
  pdf_document: default
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tlars
**Title**: The Terminated-LARS (T-LARS) algorithm for high-dimensional early terminated forward variable selection

**Description**: It computes the solution path of the Terminating-LARS (T-LARS) algorithm. The T-LARS algorithm appends dummy predictors to the original predictor matrix and terminates the forward-selection process after a pre-defined number of dummy variables has been selected.

**Note**: The T-LARS algorithm is a major building block of the T-Rex Selector ([Paper](https://arxiv.org/abs/2110.06048) and [R package](https://github.com/jasinmachkour/tknock)). The T-Rex Selector performs terminated-random experiments (T-Rex) using the T-LARS algorithm and fuses the selected active sets of each random experiment to obtain a final set of selected variables, while provably controlling the FDR at the user-defined target level and maximizing the number of selected variables.

In the following, we want to show you how to use the package and give you an idea of why terminating the solution path early is a reasonable approach in high-dimensional and sparse variable selection: In many applications, most active variables enter the solution path early!

## Installation
You can install the tlars package from [GitHub](https://github.com/jasinmachkour/tlars) with 

``` r
install.packages("devtools")
devtools::install_github("jasinmachkour/tlars")
```

You can open the help pages with
```{r, eval=FALSE}
library(tlars)
help(package = "tlars")
?tlars
?tlars_model
?tlars_cpp
?plot.Rcpp_tlars_cpp
?print.Rcpp_tlars_cpp
?Gauss_data
```

To cite the package ‘tlars’ in publications use:
```{r, eval=FALSE}
citation("tlars")
```


## Quick Start
In the following, we illustrate the basic usage of the tlars package to perform variable selection in sparse and high-dimensional regression settings using the T-LARS algorithm.

1. **First**, we generate a high-dimensional Gaussian data set with sparse support:

```{r}
library(tlars)

# Setup
n <- 75 # Number of observations
p <- 150 # Number of variables
num_act <- 3 # Number of true active variables
beta <- c(rep(1, times = num_act), rep(0, times = p - num_act)) # Coefficient vector
actives <- which(beta > 0) # Indices of true active variables
num_dummies <- p # Number of dummy predictors (or dummies)

# Generate Gaussian data
set.seed(123)
X <- matrix(stats::rnorm(n * p), nrow = n, ncol = p)
y <- X %*% beta + stats::rnorm(n)
```

2. **Second**, we generate a dummy matrix containing n rows and num_dummies dummy predictors that are sampled from the standard normal distribution and append it to the original predictor matrix:

```{r}
set.seed(1234)
dummies <- matrix(stats::rnorm(n * num_dummies), nrow = n, ncol = num_dummies)
XD <- cbind(X, dummies)
```

3. **Third**, we generate an object of the C++ class tlars_cpp and supply the information that the last num_dummies predictors in XD are dummy predictors:

```{r}
mod_tlars <- tlars_model(X = XD, y = y, num_dummies = num_dummies)
```

4. **Finally**, we perform three T-LARS-steps on "mod_tlars", i.e., the T-LARS algorithm is run until **T_stop = 3** dummies have entered the solution path and stops there. For comparison, we also compute the whole solution path by setting early_stop = FALSE:

4.1. Perform three T-LARS-steps on object "mod_tlars":

```{r terminated_solution_path, echo=TRUE, fig.align='center', message=TRUE, fig.width = 10, fig.height = 6, out.width = "90%"}
tlars(model = mod_tlars, T_stop = 3, early_stop = TRUE) # Perform three T-LARS-steps on object "mod_tlars"
print(mod_tlars) # Print information about the results of the performed T-LARS-steps
plot(mod_tlars) # Plot the terminated solution path
```

4.2. Compute the whole solution path:

```{r full_solution_path, echo=TRUE, fig.align='center', message=TRUE, fig.width = 10, fig.height = 6, out.width = "90%"}
tlars(model = mod_tlars, early_stop = FALSE) # Compute the whole solution path
print(mod_tlars) # Print information about the results
plot(mod_tlars) # Plot the whole solution path
```

## Outlook
The T-LARS algorithm is a major building block of the T-Rex Selector ([Paper](https://arxiv.org/abs/2110.06048) and [R package](https://github.com/jasinmachkour/tknock)). The T-Rex Selector performs terminated-random experiments (T-Rex) using the T-LARS algorithm and fuses the selected active sets of each random experiment to obtain a final set of selected variables, while provably controlling the FDR at the user-defined target level and maximizing the number of selected variables. Check it out if you need a fast and FDR-controlling variable/feature selection method for large-scale high-dimensional settings!

## Documentation
For more information and some examples, please check the
[GitHub-vignette](https://github.com/jasinmachkour/tlars/blob/main/vignettes/tlars_variable_selection.Rmd).


## Links
tlars package: [GitHub-tlars](https://github.com/jasinmachkour/tlars).

README file: [GitHub-readme](https://github.com/jasinmachkour/tlars/blob/main/README.html).

Vignette: [GitHub-vignette](https://github.com/jasinmachkour/tlars/blob/main/vignettes/tlars_variable_selection.Rmd).

trex package: [GitHub-trex](https://github.com/jasinmachkour/tknock).

T-Rex paper: https://arxiv.org/abs/2110.06048
