---
title: "The Terminating-LARS (T-LARS) Method: Usage and Simulations"
author: |
  | Jasin Machkour^\#^, Simon Tien^\#^, Daniel P. Palomar^\*^, Michael Muma^\#^
  |
  | ^\#^Technische Universität Darmstadt
  | ^\*^The Hong Kong University of Science and Technology
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: flatly
    highlight: pygments
    toc: yes
    toc_depth: 1
    toc_float: yes
  prettydoc::html_pretty:
    theme: tactile
    highlight: vignette
    toc: yes
    toc_depth: 2
toc-title: "Table of Contents"    
csl: the-annals-of-statistics.csl # https://www.zotero.org/styles
bibliography: refs.bib
nocite: |
  @machkour2021terminating
vignette: |
  %\VignetteKeyword{T-LARS, T-Rex selector, false discovery rate (FDR) control, high-dimensional variable selection, martingale theory, genome-wide association studies}
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{The Terminating-LARS (T-LARS) Method: Usage and Simulations}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.retina = 2,
  out.width = "85%",
  dpi = 96
  # pngquant = "--speed=1"
)
options(width = 80)
```


-----------
# Motivation
This package implements the **Terminating-LARS (T-LARS)** algorithm, i.e., it computes the solution path of the T-LARS algorithm. The T-LARS algorithm appends dummy predictors to the original predictor matrix and terminates the forward-selection process after a pre-defined number of dummy variables has been selected.

The T-LARS algorithm is a major building block of the **T-Rex Selector** ([Paper](https://arxiv.org/abs/2110.06048) and [R package](https://github.com/jasinmachkour/tknock)). The T-Rex Selector performs terminated-random experiments (T-Rex) using the T-LARS algorithm and fuses the selected active sets of each random experiment to obtain a final set of selected variables, while provably controlling the FDR at the user-defined target level and maximizing the number of selected variables.

In the following, we want to show you how to use the package and give you an idea of why terminating the solution path early is a reasonable approach in high-dimensional and sparse variable selection: In many applications, most active variables enter the solution path early!

<!-- Include Latex macros -->
$$
\DeclareMathOperator{\FDP}{FDP}
\DeclareMathOperator{\FDR}{FDR}
\DeclareMathOperator{\TPP}{TPP}
\DeclareMathOperator{\TPR}{TPR}
\newcommand{\A}{\mathcal{A}}
\newcommand{\coloneqq}{\mathrel{\vcenter{:}}=}
$$

# Installation
You can install the tlars package from [GitHub](https://github.com/jasinmachkour/tlars) with 

``` r
install.packages("devtools")
devtools::install_github("jasinmachkour/tlars")
```

You can open the help pages with
```{r, eval=FALSE}
library(tlars)
help(package = "tlars")
?tlars
?tlars_model
?tlars_cpp
?plot.Rcpp_tlars_cpp
?print.Rcpp_tlars_cpp
?Gauss_data
```

To cite the package ‘tlars’ in publications use:
```{r, eval=FALSE}
citation("tlars")
```


## Quick Start
In the following, we illustrate the basic usage of the tlars package to perform variable selection in sparse and high-dimensional regression settings using the T-LARS algorithm.

1. **First**, we generate a high-dimensional Gaussian data set with sparse support:
```{r}
library(tlars)

# Setup
n <- 75 # Number of observations
p <- 150 # Number of variables
num_act <- 3 # Number of true active variables
beta <- c(rep(1, times = num_act), rep(0, times = p - num_act)) # Coefficient vector
actives <- which(beta > 0) # Indices of true active variables
num_dummies <- p # Number of dummy predictors (or dummies)

# Generate Gaussian data
set.seed(123)
X <- matrix(stats::rnorm(n * p), nrow = n, ncol = p)
y <- X %*% beta + stats::rnorm(n)
```

2. **Second**, we generate a dummy matrix containing n rows and num_dummies dummy predictors that are sampled from the standard normal distribution and append it to the original predictor matrix:
```{r}
set.seed(1234)
dummies <- matrix(stats::rnorm(n * num_dummies), nrow = n, ncol = num_dummies)
XD <- cbind(X, dummies)
```

3. **Third**, we generate an object of the C++ class tlars_cpp and supply the information that the last num_dummies predictors in XD are dummy predictors:
```{r}
mod_tlars <- tlars_model(X = XD, y = y, num_dummies = num_dummies)
```

4. **Finally**, we perform one T-LARS-step on "mod_tlars", i.e., the T-LARS algorithm is run until **T_stop = 1** dummy has entered the solution path and stops there:

4.1. Perform one T-LARS-step on object "mod_tlars":
```{r terminated_solution_path_T_3, echo=TRUE, fig.align='center', message=TRUE, fig.width = 10, fig.height = 6, out.width = "90%"}
tlars(model = mod_tlars, T_stop = 1, early_stop = TRUE) # Perform one T-LARS-step on object "mod_tlars"
print(mod_tlars) # Print information about the results of the performed T-LARS-steps
plot(mod_tlars) # Plot the terminated solution path
```


# T-LARS Warm Starts
The object "mod_tlars" stores the results and, therefore, allows for warm starts. That is, after performing a T-LARS-step with, e.g., "T_stop = 1", we can perform another T-LARS-step with, e.g., "T_stop = 2, 3, ...", by continuing to build the solution path from its last T-LARS-step.

******

* **T_stop = 2**:

```{r terminated_solution_path_T_5, echo=TRUE, fig.align='center', message=TRUE, fig.width = 10, fig.height = 6, out.width = "90%"}
# Numerical zero
eps <- .Machine$double.eps

# Perform one additional T-LARS-step (going from T_stop = 1 to T_stop = 2) on object "mod_tlars"
tlars(model = mod_tlars, T_stop = 2, early_stop = TRUE)
print(mod_tlars)
plot(mod_tlars)

# Coefficient vector corresponding to original and dummy variables at the terminal T-LARS step 
beta_hat <- mod_tlars$get_beta()

selected_var <- which(abs(beta_hat[seq(p)]) > eps) # Indices of selected original variables
selected_dummy <- p + which(abs(beta_hat[seq(p + 1, ncol(XD))]) > eps) # Indices of selected dummy variables

FDP <- length(setdiff(selected_var, actives)) / max(1, length(selected_var)) # False discovery proportion (FDP)
TPP <- length(intersect(selected_var, actives)) / max(1, length(actives)) # True positive proportion (TPP)

selected_var
selected_dummy
FDP
TPP
```

******

* **T_stop = 5**:

```{r terminated_solution_path_T_10, echo=TRUE, fig.align='center', message=TRUE, fig.width = 10, fig.height = 6, out.width = "90%"}
# Perform three additional T-LARS-steps (going from T_stop = 2 to T_stop = 5) on object "mod_tlars"
tlars(model = mod_tlars, T_stop = 5, early_stop = TRUE)
print(mod_tlars)
plot(mod_tlars)

# Coefficient vector corresponding to original and dummy variables at the terminal T-LARS step 
beta_hat <- mod_tlars$get_beta()

selected_var <- which(abs(beta_hat[seq(p)]) > eps) # Indices of selected original variables
selected_dummy <- p + which(abs(beta_hat[seq(p + 1, ncol(XD))]) > eps) # Indices of selected dummy variables

FDP <- length(setdiff(selected_var, actives)) / max(1, length(selected_var)) # False discovery proportion (FDP)
TPP <- length(intersect(selected_var, actives)) / max(1, length(actives)) # True positive proportion (TPP)

selected_var
selected_dummy
FDP
TPP
```

******


# FDR and TPR
From a statistical point of view, it is desirable to use a variable selection method that allows for controlling the expected value of the FDP at a user-defined target level $\alpha \in [0, 1]$ while maximizing the number of selected variables. These type of methods exist and are called false discovery rate (FDR)-controlling methods. For example, the [T-Rex Selector](https://arxiv.org/abs/2110.06048) is a fast and FDR-controlling variable/feature selection framework for large-scale high-dimensional settings that relies on the T-LARS method.

**Definitions** (FDR and TPR)
Let $\widehat{\A} \subseteq \lbrace 1, \ldots, p \rbrace$ be the set of selected variables, $\A \subseteq \lbrace 1, \ldots, p \rbrace$ the set of true active variables, $| \widehat{\A} |$ the cardinality of $\widehat{\A}$, and define $1 \lor a \coloneqq \max\lbrace 1, a \rbrace$, $a \in \mathbb{R}$. Then, the false discovery rate (FDR) and the true positive rate (TPR) are defined by
$$
\FDR \coloneqq \mathbb{E} \big[ \FDP \big] \coloneqq \mathbb{E} \left[ \dfrac{\big| \widehat{\A} \backslash \A \big|}{1 \lor \big| \widehat{\A} \big|} \right]
$$
and 

$$
\TPR \coloneqq \mathbb{E} \big[ \TPP \big] \coloneqq \mathbb{E} \left[ \dfrac{| \A  \cap \widehat{\A} |}{1 \lor | \A |} \right],
$$
respectively.

# Simulations
We conduct Monte Carlo simulations and plot the resulting averaged FDP and TPP over the number of included dummies T. Note that the averaged FDP and TPP are estimates of the FDR and TPR, respectively.

```{r}
# Setup
n <- 100 # number of observations
p <- 300 # number of variables
num_act <- 10 # number of true active variables
beta <- c(rep(3, times = num_act), rep(0, times = p - num_act)) # coefficient vector
actives <- which(beta > 0) # indices of true active variables
num_dummies <- p # number of dummies
T_vec <- c(1, 2, 5, 10, 20, 50, 100) # stopping points, i.e, number of included dummies before terminating the solution path
MC <- 500 # number of Monte Carlo runs per stopping point

# Initialize results vectors
FDP <- matrix(NA, nrow = MC, ncol = length(T_vec))
TPP <- matrix(NA, nrow = MC, ncol = length(T_vec))

# Numerical zero
eps <- .Machine$double.eps

# Seed
set.seed(12345)

# Run simulations
for (t in seq_along(T_vec)) {
  for (mc in seq(MC)) {
    # Generate Gaussian data
    X <- matrix(stats::rnorm(n * p), nrow = n, ncol = p)
    y <- X %*% beta + stats::rnorm(n)

    # Generate dummy matrix and append it to X
    dummies <- matrix(stats::rnorm(n * p), nrow = n, ncol = num_dummies)
    XD <- cbind(X, dummies)

    # Create object of class tlars_cpp
    mod_tlars <- tlars_model(X = XD, y = y, num_dummies = num_dummies, type = "lar", info = FALSE)

    # Run T-LARS-steps
    tlars(model = mod_tlars, T_stop = t, early_stop = TRUE, info = FALSE)
    beta_hat <- mod_tlars$get_beta()
    selected_var <- which(abs(beta_hat[seq(p)]) > eps)

    # Results
    FDP[mc, t] <- length(setdiff(selected_var, actives)) / max(1, length(selected_var))
    TPP[mc, t] <- length(intersect(selected_var, actives)) / max(1, length(actives))
  }
}

# Compute estimates of FDR and TPR by averaging FDP and TPP over MC Monte Carlo runs
FDR <- colMeans(FDP)
TPR <- colMeans(TPP)
```

```{r FDR_and_TPR, echo=FALSE, fig.align='center', message=FALSE, fig.width = 9, fig.height = 5, out.width = "90%"}
# Plot results
par(mfrow = c(1, 2)) # 1 x 2 plot-grid

# FDP vs. T
plot(T_vec, FDR, type = "b", xlab = "T", xlim = c(0, 100), ylim = c(0, 1), lwd = 1)
grid()

# TPP vs. T
plot(T_vec, TPR, type = "b", xlab = "T", xlim = c(0, 100), ylim = c(0, 1), lwd = 1)
grid()

par(mfrow = c(1, 1)) # reset plot-grid

# TPP vs. FDP
plot(FDR, TPR, type = "b", ylim = c(0, 1), lwd = 1)
grid()
```


# Outlook
The T-LARS algorithm is a major building block of the T-Rex Selector ([Paper](https://arxiv.org/abs/2110.06048) and [R package](https://github.com/jasinmachkour/tknock)). The T-Rex Selector performs terminated-random experiments (T-Rex) using the T-LARS algorithm and fuses the selected active sets of each random experiment to obtain a final set of selected variables, while provably controlling the FDR at the user-defined target level and maximizing the number of selected variables. Check it out if you need a fast and FDR-controlling variable/feature selection method for large-scale high-dimensional settings!

# References {-}
